# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r Loading and preprocessing the data}
library(ggplot2)
csv=read.csv("activity.csv", na.strings=c("NA"))
csv$date<-as.Date(csv$date, format="%Y-%m-%d")
```

## What is mean total number of steps taken per day?

```{r histogram total steps per day, message=FALSE}
sumByDate<-aggregate(csv$steps,by=list(csv$date),sum)
sumByDateNoNA<-na.omit(sumByDate)
names(sumByDateNoNA)<-c("date","steps")
qplot(sumByDateNoNA$steps, data=sumByDateNoNA, xlab="Total Steps")

themean<-mean(sumByDateNoNA$steps)
themedian<-median(sumByDateNoNA$steps)
```

The mean is `r themean`

The median is `r themedian`

## What is the average daily activity pattern?

```{r time series plot of averaged steps}
avgStepsPerInterval<-aggregate(steps~interval,data=csv,FUN="mean")
qplot(x=interval, y=steps, geom="line", data=avgStepsPerInterval)

max<-avgStepsPerInterval[avgStepsPerInterval$steps==max(avgStepsPerInterval$steps),]

```
The Interval containing the maximum number of steps is `r max$interval`

## Imputing missing values
### Strategy: I will replace the mean for that 5-minute interval for the NAs

```{r imputed histogram, message=FALSE}
totalNumberOfMissingRows<-nrow(csv)-nrow(csv[complete.cases(csv),])
replaced<-csv
for (i in 1:nrow(replaced)) {
        if (is.na(replaced$steps[i])) {
                replaced$steps[i]<-avgStepsPerInterval$steps[avgStepsPerInterval$interval==replaced$interval[i]]
        }
}


replacedDataSumByDate<-aggregate(replaced$steps,by=list(replaced$date),sum)
names(replacedDataSumByDate)<-c("date","steps")

qplot(replacedDataSumByDate$steps, data=replacedDataSumByDate, xlab="Total Steps")
replacedMean<-mean(replacedDataSumByDate$steps)
replacedMedian<-median(replacedDataSumByDate$steps)
```

Total number of missing rows is `r totalNumberOfMissingRows`

The new mean is `r replacedMean`

The new median is `r replacedMedian`

Imputing missing data did not change mean but changed median. Looking at the previous plot it looks like it didnt make much difference

## Are there differences in activity patterns between weekdays and weekends?

```{r load libs, include=FALSE, message=FALSE}
install.packages("gridExtra", repos="http://cran.rstudio.com/")
library(lattice)
library(gridExtra)
```

```{r weekday weekend plot}
# Introduce factor weekday or weekend
replaced$type<-factor(ifelse((weekdays(replaced$date)=="Saturday" | weekdays(replaced$date)=="Sunday"), "weekend", "weekday"))

replacedWE<-subset(replaced, type=="weekend")
replacedWD<-subset(replaced, type=="weekday")

replacedAvgWD<-aggregate(steps~interval,replacedWD,mean)
replacedAvgWE<-aggregate(steps~interval,replacedWE,mean)

p1<-xyplot(steps~interval, data=replacedAvgWE, type="l",main="weekend", ylab="Number of steps")
p2<-xyplot(steps~interval, data=replacedAvgWD, type="l",main="weekday", ylab="Number of steps")


grid.arrange(p1,p2,ncol=1)
```